---
title: "Descriptive Statistics"
author: "Amina Lampkin"
date: "10/13/2020"
output: html_document
---
# 
```{r echo=FALSE}
# Import libraries
library(dplyr)
library(readr)

# Open the demographic and loan file
dl = read.csv(file = "/Users/aminasymone/Documents/JOUR355/CleanedFiles/loan_and_demo_moe_clean_102020.csv",
              header = TRUE,
              sep = ",")
demo = read.csv(file = "/Users/aminasymone/Documents/JOUR355/CleanedFiles/clean_demographics_101520.csv",
                header = TRUE,
                sep = ",")

# Select the columns of interest for the demographic
demo = select(demo, zip, estimate_pop, pct_white, pct_black, est_med_inc)

```
# Zip Codes by Loan Amount

```{r}
# Aggregate by zip code for avg loan amount
avg_loans <- aggregate(x = dl$LoanAmount,
                       by = list(dl$zip),
                       FUN = mean)

# Sort by descending order
avg_loans = arrange(avg_loans, desc(x))

# Rename the columns and join the demographic data
avg_loans = rename(avg_loans, avg_loan = x)
avg_loans = rename(avg_loans, zip = Group.1)

# Create a file for the aggregated columns
write.csv(avg_loans, file = "/Users/aminasymone/Documents/JOUR355/Analysis/entire_zip_avgs.csv")


avg_loans = left_join(avg_loans, 
                      demo,
                      by = "zip")

# Pull the top and bottom 10 statistics
write.csv(head(avg_loans, 10), file = "/Users/aminasymone/Documents/JOUR355/Analysis/avg_top_10.csv" )
write.csv(tail(avg_loans, 10), file = "/Users/aminasymone/Documents/JOUR355/Analysis/avg_bottom_10.csv" )

```

# Predominantly White Zip Codes
In this chunk of code, we'll be pulling the zip codes with a majority white population (>50%) and finding out what funding they received. 
```{r}
# Pull the zip code records with a majority population of white people
white_zips = filter(dl, pct_white>=50.0)
write.csv(white_zips, file = "/Users/aminasymone/Documents/JOUR355/Analysis/white_zip_codes.csv")


# Find the sums of loans in each zip code
avg_white_zips = aggregate(x = white_zips$LoanAmount, 
                           by = list(white_zips$zip),
                           FUN = mean)

#Sort in descending order by average loan amount
avg_white_zips = arrange(avg_white_zips, desc(x))

# Rename the columns and join the demographic data
avg_white_zips = rename(avg_white_zips, avg_loan = x)
avg_white_zips = rename(avg_white_zips, white_zip = Group.1)

# Create a file for the aggregated columns
write.csv(avg_white_zips, file = "/Users/aminasymone/Documents/JOUR355/Analysis/white_zip_avgs.csv")

avg_white_zips = left_join(avg_white_zips, 
                      demo,
                      by = c("white_zip" = "zip"))
write.csv(white_zips, file = "/Users/aminasymone/Documents/JOUR355/Analysis/white_zip_codes.csv")

# Pull the top and bottom 10 statistics
write.csv(head(avg_white_zips, 10), file = "/Users/aminasymone/Documents/JOUR355/Analysis/avg_white_top_10.csv" )
write.csv(tail(avg_white_zips, 10), file = "/Users/aminasymone/Documents/JOUR355/Analysis/avg_white_bottom_10.csv" )

```

# Predominantly Black Zip Codes
In this chunk of code, we'll be pulling the zip codes with a majority Black population (>50%) and finding out what funding they received. 

```{r}
# Pull the zip code records with a majority population of Black people
black_zips = filter(dl, pct_black>=50.0)
write.csv(black_zips, file = "/Users/aminasymone/Documents/JOUR355/Analysis/black_zip_codes.csv")


# Find the sums of loans in each zip code
avg_black_zips = aggregate(x = black_zips$LoanAmount, 
                           by = list(black_zips$zip),
                           FUN = mean)

#Sort in descending order by average loan amount
avg_black_zips = arrange(avg_black_zips, desc(x))

# Rename the columns and join the demographic data
avg_black_zips = rename(avg_black_zips, avg_loan = x)
avg_black_zips = rename(avg_black_zips, black_zip = Group.1)

# Create a file for the aggregated columns
write.csv(avg_black_zips, file = "/Users/aminasymone/Documents/JOUR355/Analysis/black_zip_avgs.csv")

avg_black_zips = left_join(avg_black_zips, 
                      demo,
                      by = c("black_zip" = "zip"))
write.csv(black_zips, file = "/Users/aminasymone/Documents/JOUR355/Analysis/black_zip_codes.csv")

# Pull the top and bottom 10 statistics
write.csv(head(avg_black_zips, 10), file = "/Users/aminasymone/Documents/JOUR355/Analysis/avg_black_top_10.csv" )
write.csv(tail(avg_black_zips, 10), file = "/Users/aminasymone/Documents/JOUR355/Analysis/avg_black_bottom_10.csv" )

```

## Majority minority zip codes
In this chunk of code, we'll be pulling the zip codes with a majority minority population (percent white < 50%) and finding out what funding they received. 

```{r}
# Pull the zip code records with a majority population of Black people
minority_zips = filter(dl, pct_white < 50.0)
write.csv(minority_zips, file = "/Users/aminasymone/Documents/JOUR355/Analysis/minority_zips.csv")


# Find the sums of loans in each zip code
avg_minority_zips = aggregate(x = minority_zips$LoanAmount, 
                           by = list(minority_zips$zip),
                           FUN = mean)

#Sort in descending order by average loan amount
avg_minority_zips = arrange(avg_minority_zips, desc(x))

# Rename the columns and join the demographic data
avg_minority_zips = rename(avg_minority_zips, avg_loan = x)
avg_minority_zips = rename(avg_minority_zips, min_zip = Group.1)

# Create a file for the aggregated columns
write.csv(avg_minority_zips, file = "/Users/aminasymone/Documents/JOUR355/Analysis/minority_zip_avgs.csv")


avg_minority_zips = left_join(avg_minority_zips, 
                      demo,
                      by = c("min_zip" = "zip"))
write.csv(minority_zips, file = "/Users/aminasymone/Documents/JOUR355/Analysis/minority_zips.csv")


# Pull the top and bottom 10 statistics
write.csv(head(avg_minority_zips, 10), file = "/Users/aminasymone/Documents/JOUR355/Analysis/avg_minority_top_10.csv" )
write.csv(tail(avg_minority_zips, 10), file = "/Users/aminasymone/Documents/JOUR355/Analysis/avg_minority_bottom_10.csv" )

```

# Summary Stats
In this chunk of code we will be developing the summary statistics for the entire dataset, the Black zip codes, and minority zip codes.  
```{r}
library(ggplot2)

## Entire Dataset
summary(dl)
### Top and Bottom 10 Loans
arrange(dl, desc(dl$LoanAmount))
write.csv(head(dl, 10), file = "/Users/aminasymone/Documents/JOUR355/Analysis/top_10_loans.csv")
write.csv(tail(dl, 10), file = "/Users/aminasymone/Documents/JOUR355/Analysis/bottom_10_loans.csv")

## Black zip codes
summary(black_zips)
### Top and Bottom 10 Loans
arrange(black_zips, desc(black_zips$LoanAmount))
write.csv(head(black_zips, 10), file = "/Users/aminasymone/Documents/JOUR355/Analysis/black_zips_top_10_loans.csv")
write.csv(tail(black_zips, 10), file = "/Users/aminasymone/Documents/JOUR355/Analysis/black_zips_bottom_10_loans.csv")

## Minority zip codes
summary(minority_zips)
### Top and Bottom 10 Loans
arrange(minority_zips, desc(minority_zips$LoanAmount))
write.csv(head(minority_zips, 10), file = "/Users/aminasymone/Documents/JOUR355/Analysis/minority_zips_top_10_loans.csv")
write.csv(tail(minority_zips, 10), file = "/Users/aminasymone/Documents/JOUR355/Analysis/minority_zips_bottom_10_loans.csv")


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
