---
title: "ACS Variable Exploration"
author: "Amina Lampkin"
date: "10/1/2020"
output: html_document
---
# Prepare: get the packages and data 
```{r}

# Tidycensus and Tidyverse
library(tidycensus)
library(tidyverse)

# Define the api key
census_api_key("c1bfa3954f446908024748a043f763beee1a2069")

# View the variables for the American Community Survey
acs_variable_list <- load_variables(2018, "acs5", cache = FALSE)
```

# Variables of interest (by zip code)
- Total population (from RACE): B02001_001
- Black (alone) population by zip code: B02001_003 (total)
- White population by zip code: B02001_002
- Sex Total (Male): B01001_002	
- Sex Total (Female): B01001_026
- Median income: B19326_001 (2018 inflation-adjusted)
- Income in the past 12 months below poverty level: B17001_002 (total) (M: B17001_003) (F: B17001_017)

# Calculations for each
##  Make a list of all the variables you make, do a for loop that goes through each and calculates pct_est + pct_moe
- Percent version of the estimate
- Margin of error percentage (should be less than 10, shooting for less than 5) 

*ZCTAs can only be requested for the entire country (not within states or counties) -- need a way to filter out a list of all the MD ZCTAs -- join this document by GEOID on the PPP data and drop off the GEOIDs that don't match (inner join)* 

## Total Population
```{r}
# Get ACS total population
race_pop <- get_acs(variables = "B02001_001",
                     geography = "zcta",
                     geometry = FALSE)%>%
  mutate(pct_moe = moe/estimate*100)

```

## Black (alone) population by zip code: B02001_003 (total)
```{r}
# Get Black (alone) population
zcta_black <-get_acs(variables = "B02001_003",
                     geography = "zcta",
                     summary_var= "B02001_001",
                     geometry = FALSE)%>%
  mutate(pct_moe_black = (moe/estimate)*100)%>%
  mutate(pct_black = (estimate/summary_est)*100)
```

## White population by zip code: B02001_002
```{r}
# Get white (alone) population
zcta_white <- get_acs(variables ="B02001_002",
                      geography = "zcta",
                      summary_var = "B02001_001",
                      geometry= FALSE)%>%
    mutate(pct_moe_white = (moe/estimate)*100)%>%
    mutate(pct_white = (estimate/summary_est)*100)


```
## Sex Total (Male): B01001_002
```{r}
# Get the male population
zcta_male <- get_acs(variables ="B01001_002",
                      geography = "zcta",
                      summary_var = "B02001_001",
                      geometry= FALSE)%>%
  mutate(pct_moe_male = (moe/estimate)*100)%>%
  mutate(pct_male = (estimate/summary_est)*100)


```
## Sex Total (Famale): B01001_026
```{r}
# Get the female population
zcta_female <- get_acs(variables ="B01001_026",
                      geography = "zcta",
                      summary_var = "B02001_001",
                      geometry= FALSE)%>%
    mutate(pct_moe_female = (moe/estimate)*100)%>%
    mutate(pct_female = (estimate/summary_est)*100)


```

# Median income: B19326_001 (2018 inflation-adjusted)
```{r}
# Get the median income
zcta_median_inc <- get_acs(variables = "B19326_001",
                             geography  = "zcta",
                             geometry = FALSE)%>%
  mutate(pct_moe_med_inc = (moe/estimate)*100)
```

# Income in the past 12 months below poverty level: B17001_002 (total) 
```{r}
# Get the number of people with poverty status
zcta_poverty <- get_acs(variables = "B17001_002",
                             geography  = "zcta",
                             summary_var = "B02001_001",
                             geometry = FALSE)%>%
  mutate(pct_moe_pov = (moe/estimate)*100)%>%
  mutate(pct_poverty = (estimate/summary_est)*100)
```
# Join all the tables together to form one large table
## Will need to rename columns so I know which estimates belong to what data subset
```{r}
library(dplyr)

write.csv(race_pop, file = "race_pop.csv")
write.csv(zcta_black, file = "zcta_black.csv")
write.csv(zcta_white, file = "zcta_white.csv")
write.csv(zcta_male, file = "zcta_male.csv")
write.csv(zcta_female, file = "zcta_female.csv")
write.csv(zcta_median_inc, file = "zcta_median_inc.csv")
write.csv(zcta_poverty, file = "zcta_poverty.csv")

```